---
title: "Assignment 4"
author: "Clifford Isenor"
date: "25/02/2020"
output: pdf_document
---


### Packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(dslabs)
```

## get data
```{r}
filepath <-"https://quantdev.ssri.psu.edu/sites/qdev/files/student-mat.csv"
df <-read.table(file=url(filepath),sep=";",header=TRUE)

```

# Q1
### create a matrix including only factor variables from the data. Call this ma-trix XF. Do not do it manually. 
### You must use a loop or lapply/sapply functions.See earlier lectures. (10 Points)
```{r}

xf <-as.matrix(df[,sapply(df, is.factor)])

```

# Q2
### Create a matrix including only integer variables from the data.  Call this matrix XC. Change every integer variable in XC to numeric variable and scale them (see scale()). Do not do it manually. You must use a loop or lapply/sapplyfunctions (15 Points).

```{r}
xc <- as.matrix(df[,sapply(df, is.integer)])
xc <- scale(xc)

```

# Q3
### Model.matrix() with XF to create numeric (0 and 1) columns.  (5Points).
```{r}
xff <- model.matrix(~.-1, data.frame(xf))

```

# Q4
```{r}
y <- factor(xc[,16],levels = c(0,1), labels = c("i < Mean", "i < Mean") )#establish container of factor var
y<-ifelse((xc[, 16] < mean(xc[,16])),1,0) #assign the 1 if G# test score < mean esle 0
xc<-xc[,-(14:16)] # drop G1,G2,G3
```

# Q5
```{r}
df <- data.frame(y,xff,xc)

```

# Q6
### 1) Split data
```{r}

sh <- sample(nrow(df), nrow(df), replace = FALSE) #shuffle data
ind_test <- sh[1:(nrow(df) / 10)] #index postioins of 10% of df data
ind_train <- sh[-ind_test] #index postioins of 90% of df data

test<-df[ind_test,] #10% of df
train<- df[ind_train,] #90% of df

k<-10
```


### 2) hyperperameters
```{r}
knn <- seq(from = 1, to = 30, by = 1 ) #the numbers of grouping s to be tested
th <- seq(from = 0.045, 0.7, by = 0.025) #the thresholds to use

```

### 3) Containtiners
```{r}
hit <- matrix(0, length(th), length(knn)) #hit ratio container
opt <- matrix(0,k,3) #optimal vaules container
best <- c()
```

### 4) use KLM Model to predict y
```{r}
library(caret)

ntest <- floor(nrow(train) / k)


for (i in 1:k) {#loop for [k]number of folds
  
  ind_test <- c(((i - 1) * ntest + 1):(i * ntest)) #get index for test slice(i)
  ind_train <- c(1:nrow(train))[-ind_test] #get index for train slice(i)

  df_test <- train[ind_test,] #get actual vaules for test slice[i]
  df_train <- train[ind_train,] #get actual vaules for train slice[i]
  
  
  for (s in 1:length(knn)) { #number of groupings = knn
    
    #run model
    model <-
      knn3(y ~ .,
           data = df_train,
           k = knn[s])
    #predict
    phat <-
      predict(model, df_test, type = "prob") #breaks here on i=10 for some
    
    for (j in 1:length(th)) {
      #try all threshold vaules
      
      yhat <-
        phat[, 2] > th[j] #test predicited vaules against a seris of thresholds; True if greater then then thres else False
      
      ct <-
        table(df_test$y, yhat) #create a confusion table with results
      
      hit[j, s] <-
        sum(diag(ct) / sum(ct)) #calculate hit ratio from confusion table
      
      cat("k-fold loop: ", i, ".", s, '.', j,  "\r") #cycle counter
    }
  }
  #theres got to be a better way
  coor <-
    as.vector((which(hit == max(hit), arr.ind = TRUE))[1, ]) # coordantes of first highest hit ration
  
  opt[i, ] <-
    c(coor[1], coor[2], hit[coor[1], coor[2]]) #assign coorordante 1,#assign coorordante 2, and respective hit rate
  
}
best <- max(opt[, 3]) # find highest over all hit rate
opt #print results
cat('\n best over all: ', best) #print best

```

